---
title: "Cross-Cluster Weighted Forests"
output: github_document
vignette: >
  %\VignetteIndexEntry{Cross-Cluster Weighted Forests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

The `rCCWF` package implements Cross-Cluster Weighted Forests, an ensemble learning approach that combines clustering with random forests to improve prediction accuracy. This method is particularly effective when dealing with heterogeneous data that naturally forms clusters or when different subpopulations exist within your dataset.

Load the necessary packages:

```{r}
library(rCCWF)
library(dplyr)
library(ggplot2)
```

## Basic Usage

Simple example using simulated data:

```{r}
# Set seed for reproducibility
set.seed(123)

# Simulate data with known clusters
sim_data <- sim_data_monte(
  nclusters = 5,  # Number of training clusters
  ncoef = 20,     # Number of predictor variables
  ntest = 2       # Number of test datasets
)

# Examine the structure
names(sim_data)
```

## Working with Pre-clustered Data

If you already have data that's naturally clustered or split:

```{r}
# Use the pre-clustered data directly
results <- cluster_ensemble_workflow(
  train_data = sim_data$clusters_list,
  test_data = sim_data$test_list,
  cluster_ind = FALSE,  # Don't perform additional clustering
  merged_trees = 500,
  cluster_trees = 100
)
```

## Using k-means clustering on data that is partitioned

If you already have data that's naturally clustered or split:

```{r}
# Use the pre-clustered data directly
results <- cluster_ensemble_workflow(
  train_data = sim_data$clusters_list,
  test_data = sim_data$test_list,
  cluster_ind = TRUE,  # We do perform additional clustering
  n_clusters = 5,
  merged_trees = 500,
  cluster_trees = 100
)

print(results$average_performance)
```

## Understanding the Predictions

The package provides four types of predictions:

1. `merged`: Predictions from a single random forest trained on all data
2. `unweighted`: Simple average of individual cluster model predictions
3. `stack_ridge`: Ridge regression weighted combination
4. `stack_lasso`: Lasso regression weighted combination

```{r}
# Examine predictions for first test set
head(results$predictions[[1]])
```

## Performance Comparison

```{r}
# For multiple test sets
print(results$average_performance)

# Visualize performance
plot_performance_comparison(results)
```

## Working with Single Datasets

For a single dataset that needs to be clustered (set cluster_ind = TRUE in this case):

```{r}
# Combine training data
combined_train <- sim_data$clusters_list |> dplyr::bind_rows()
test_data <- sim_data$test_list[[1]]

# Run workflow with automatic clustering
results_single <- cluster_ensemble_workflow(
  train_data = combined_train,
  test_data = test_data,
  n_clusters = 5,
  cluster_ind = TRUE,
  merged_trees = 500,
  cluster_trees = 100
)

# Plot performance comparison
plot_performance_comparison(results_single)
```


## Advanced Usage: Simulation Studies

The package includes functions for simulation studies:

```{r}
# Compare Gaussian and non-Gaussian clusters
gaussian_data <- sim_data_gaussian(5, 20, 2)
monte_data <- sim_data_monte(5, 20, 2)

# Run analysis on both
results_gaussian <- cluster_ensemble_workflow(
  train_data = gaussian_data$clusters_list,
  test_data = gaussian_data$test_list,
  cluster_ind = FALSE
)

results_monte <- cluster_ensemble_workflow(
  train_data = monte_data$clusters_list,
  test_data = monte_data$test_list,
  cluster_ind = FALSE
)

# Compare results
bind_rows(
  Gaussian = results_gaussian$average_performance,
  Monte = results_monte$average_performance,
  .id = "simulation"
) |>
  arrange(method)
```

## Tips and Best Practices

1. **Choosing the Number of Clusters**
   - Start with domain knowledge if available
   - Try different values and compare performance
   - Consider the silhouette score or other clustering metrics

2. **Tree Parameters**
   - Use more trees (`merged_trees`) for the merged model
   - Fewer trees (`cluster_trees`) for individual cluster models
   - Balance computation time vs. performance

3. **When to Use CCWF**
   - Heterogeneous data with natural clusters, multiple datasets
   - I.e. large datasets with potential subpopulations, data prone to batch effects, etc.

## Session Info

```{r}
sessionInfo()
```
```

